\documentclass[runningheads,a4paper]{llncs}

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[applemac]{inputenx} % This is required to support characters in spanish
\usepackage[T1]{fontenc} % This is required to have small-bold-capitals
\usepackage{fancyhdr} % This is required to have nice headers and footers
\usepackage{framed} % This is required to have framed comments
\usepackage{color} % This defines colors that can be used in the text
\usepackage{soul} % This allows the usage of \hl to highlight text
\usepackage{tocloft} % This is required to have custom lists of .....
\usepackage{comment} % This is required to remove specific environments
\usepackage{caption}
\usepackage{ifthen}
\usepackage{array}
\usepackage{wrapfig}
\usepackage{hyperref}
\usepackage{pifont}
\usepackage{listings}
\usepackage{syntax}
\usepackage{graphicx}
\usepackage{url}
\usepackage{latexsym}
\usepackage{fancybox}
\usepackage{pifont}
\usepackage{lineno}
\usepackage{todonotes}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{xspace}
\usepackage{setspace}
\usepackage{todonotes}
\usepackage{bbding}
\usepackage{xcolor,colortbl}

\lstset{ %
language=C++,                % choose the language of the code
basicstyle=\footnotesize,       % the size of the fonts that are used for the code
numbers=left,                   % where to put the line-numbers
numberstyle=\footnotesize,      % the size of the fonts that are used for the line-numbers
stepnumber=1,                   % the step between two line-numbers. If it is 1 each line will be numbered
numbersep=5pt,                  % how far the line-numbers are from the code
backgroundcolor=\color{white},  % choose the background color. You must add \usepackage{color}
showspaces=false,               % show spaces adding particular underscores
showstringspaces=false,         % underline spaces within strings
morekeywords={language,syntax,with},
showtabs=false,                 % show tabs within strings adding particular underscores
frame=single,           % adds a frame around the code
tabsize=2,          % sets default tabsize to 2 spaces
captionpos=b,           % sets the caption-position to bottom
breaklines=true,        % sets automatic line breaking
breakatwhitespace=false,    % sets if automatic breaks should only happen at whitespace
escapeinside={\%*}{*)}          % if you want to add a comment within your code
}

\setcounter{tocdepth}{3}
\newtheorem{mydef}{Definition}


\newcommand{\mc}[2]{\multicolumn{#1}{c}{#2}}
\definecolor{Gray}{gray}{0.9}
\definecolor{LightGray}{gray}{0.98}

\newcolumntype{a}{>{\columncolor{Gray}}c}
\newcolumntype{b}{>{\columncolor{white}}c}

\newcommand\etal[0]{\emph{et al.}\xspace}
\newcommand\td[1]{\todo[inline]{#1}\xspace}
\newcommand\toolname[0]{Puzzle\xspace}

\setcounter{tocdepth}{3}


\urldef{\mailsa}\path|{david.mendez-acuna,jagalindo,benoit.combemale,arnaud.blouin,benoit.baudry}@inria.fr|    
\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

\begin{document}

\mainmatter  % start of an individual contribution

\title{Identifying Reusable Language Modules from Existing Domain-Specific Languages}
\titlerunning{Identifying Reusable Language Modules from Existing DSLs}


\author{David M\'endez-Acu\~na \and Jos\'e A. Galindo \and Benoit Combemale \and \\ Arnaud Blouin \and Benoit Baudry}
\institute{University of Rennes 1, INRIA/IRISA. France\\
\vspace{1mm}\mailsa}
\authorrunning{David M\'endez-Acu\~na et. al}

\maketitle

\begin{abstract} 

The use of domain-specific languages (DSLs) has become a successful technique in the implementation of complex systems. However, the construction of this type of languages is time-consuming and requires highly-specialized knowledge and skills. 
In such context, a common practice for enabling reuse is the definition of languages modules that can be later put together  in order to build up new DSLs. Nevertheless, the identification and reuse of those modules in a set of existing languages is currently relying on manual and costly tasks thus, hindering the reuse exploitation when developing DSLs. 
%Hence, researchers are currently seeking approaches to leverage reuse during the DSLs development in order to minimize implementation from scratch. Indeed, we can find several approaches that support modular definition of DSLs. Still, the design of language modules that can be actually reusable is a difficult task. With little information, language designers have to define language modules intended to be reused in future contexts that, of course, they do not know in advance. 
In this paper, we propose a solution to i) detect and identify possible reuse in a set of DSLs; and ii) break down those languages into modules that can be later exploited to minimize costs when constructing new languages and when maintaining the existing ones. Finally, we validate our approach by using a realistic DSLs coming out from industrial projects and obtained from public \texttt{GitHub} repositories.

\end{abstract}

\input{sections/introduction}
\input{sections/basicconcepts}
\input{sections/runningexample}
%\input{sections/motivation}
\input{sections/familiesanalysis}
\input{sections/casestudies}
\input{sections/puzzlemetrics}
%\input{sections/threadstovalidity}
\input{sections/relatedwork}
\input{sections/conclusions}

\section*{Acknowledgments}
The research presented in this paper is supported by the European Union within the FP7 Marie Curie Initial Training Network ``RELATE" under grant agreement number 264840; the VaryMDE project, a collaboration between Thales and INRIA; and the Cybersecurity project involving DGA (French department of defense) and INRIA. We also thank the support provided by the Grid5000 infrastructure during the execution of our experiments. 

% BIBLIOGRAPHY
\bibliographystyle{abbrv}
\bibliography{0-main}

\end{document}
\grid
